--
Name: Zanub Hassan
title: "IsyE 603 HW 5 "
output: html_notebook
Date: March 28, 2022
---

## Problem 1
a) MA(3) - The ACF has three spikes and the PACF dies out in an exponential manner.

b) ARIMA(2,0,1)

c) ARIMA(1,1,1)

d)AR(2) - This is because the ACF plot shows a sinusodial pattern, while the PACF shows just two significant spikes


```{r}
#Problem 2(c)
#Using arima.sim function in R to simulate the three models and plot the simulated series. Make the length of the time series be 100.
library(graphics)
#first model
i<-arima.sim(model = list(order = c(1,0,1), ar = 0.7, ma = 0.6), n = 100)
plot(i)
acf(i)
pacf(i)

#second model
ii<-arima.sim(model = list(order = c(1,1, 1), ar = 0.7, ma = 0.6), n = 100)
plot(ii)
acf(ii)
pacf(ii)

#third model
 library(forecast)
    model_1 <- Arima(ts(rnorm(100),freq=4), order=c(1,0,1), seasonal=c(0,1,0),
                 fixed=c(phi=0.7, theta=-0.6))
    foo <- simulate(model_1, nsim=1000)
    plot(foo)
    Acf(foo)
    pacf(foo)
```


```{r}
#Problem 3a
#Getting the ARIMA Model
library(fpp3)
fit <- aus_airpassengers %>%
  model(ARIMA(Passengers))
report(fit)

#Plotting the ACF and PACF of residuals
fit %>% gg_tsresiduals()
#This code is supposed to plot the PACF alongside the ACF but it comes in as just a straight line 
gg_tsdisplay(residuals(fit))

#Plotting the forecasts for the next 10 periods
fit %>% forecast(h=10) %>%
  autoplot(aus_airpassengers)

#Problem 3c
#Plotting theforecast of a model from (0,1,0) with drift
fit1 <- aus_airpassengers %>%
model(arima = ARIMA(Passengers ~ pdq(0, 1, 0)))  

#plot forecast for next 10 periods
fit1 %>% forecast(h=10) %>%
  autoplot(aus_airpassengers)

#Problem 3d
#Plotting theforecast of a model from (2,1,2) with drift
fit2 <- aus_airpassengers %>%
# 1 is the drift
model(arima = ARIMA(Passengers ~ 1 + pdq(2,1,2)))  

#plot forecast for next 10 periods
fit2 %>% forecast(h=10) %>%
  autoplot(aus_airpassengers)
```

## Problem 3a
The model is ARIMA(0,2,1)

## Problem 3b
Note: E is epsilon in this case.
The model can be written as follows:
Yt = -0.8963Et-1 + Et = (-0.8963B + 1)Et

## Problem 3c
The model ARIMA (0,1,0) with the drift has an upward trend like the model in part a . However, the trend is not as rapid as that of part a.

## Problem 3d
The model ARIMA (2,1,2) with the drift has an upward trend like the model in part a . However, the trend is not as rapid as that of part a.

```{r}
#Problem 4a
library(zoo)
library(forecast)
SouvenirSales.data <- read.csv(file ='SouvenirSales.csv', header = TRUE)
#creating the data in a time series format spanning from  Jan 1995 to  Dec. 2001 using a freq of 12
Sales.ts <- ts(SouvenirSales.data$Sales, start = c(1995,1), end = c(2001, 12), freq = 12)
fixed.nValid <- 12
# since we have a fixed valid the rest will be training. 
fixed.nTrain <- length(Sales.ts) - fixed.nValwid
#training period
train.ts <- window(Sales.ts, start = c(1995, 1), end = c(1995, fixed.nTrain))
#validation period
valid.ts <- window(Sales.ts, start = c(1995, fixed.nTrain + 1), end = c(1995, fixed.nTrain + fixed.nValid))

##Running the regression for modelA- used only training period only
Model_A <- tslm(train.ts ~ trend + season)
summary(Model_A)

#Problem 4b
#Plot ACF and PACF plot until lag 10 for residuals model in part (a)
Acf(Model_A$residuals, lag.max = 10, main = "")
Pacf(Model_A$residuals, lag.max = 10, main = "")

#Problem 4c
#Establish an AR(2) model for the residuals in (b)
AR_Model <- Arima(Model_A$residuals, order =c(2,0,0))  
AR_Model

#Problem 4d
#Plot ACF and PACF plot until lag 10 for residuals model for the AR(2) model
Acf(AR_Model$residuals, lag.max = 10, main = "")
Pacf(AR_Model$residuals, lag.max = 10, main = "")

#Problem 4e
#Use auto.arima function to automatically find a best ARIMA model for the training period data
Model<-auto.arima(train.ts, seasonal=TRUE)
summary(Model)

#Problem 4f
#Fit the identified model in (e) and plot the fitted values of the model and the observed training period data.
fit4 <- Arima(train.ts, order=c(0,1,2), seasonal=c(0,1,1))
summary(fit4)

plot(train.ts, ylim = c(1500, 105000), 