---
Name: "Zanub Hassan"
title: "IsyE 603 HW1"
output: html_notebook

---


```{r}
library(tidyverse)
library(ggplot2)
#Problem 1 1
college <- read.csv(file = 'College.csv', header = TRUE)
fix(college)
rownames(college) = college[,1]
fix(college)
college = college[,-1]
fix(college)

summary(college)
# Making the pairs function read the first column as numeric data
college[,1] = as.numeric(factor(college[,1]))
pairs(college[, 1:10])

boxplot(Outstate~Private, data = college)

Elite = rep("No", nrow(college))
Elite[college$Top10perc > 50] = "Yes"
Elite = as.factor(Elite)
college = data.frame(college, Elite)
summary(college)

boxplot(Outstate~Elite, data = college)
# The code directly below will partition  the print window into 4 to enable 4 plots to be made simultaneously
par(mfrow = c(2,2))
hist(college$Top10perc, breaks = 5)
hist(college$Top10perc, breaks = 10)
hist(college$Top10perc, breaks = 20)
hist(college$Top10perc, breaks = 30)

ggplot(data = college, aes(Apps, Accept)) +
  geom_point(shape = 21, alpha= 0.3)+
  geom_smooth(method = lm)

```
## Problem 1(vi)
There is a positive correlation between the number of applications and the amount of acceptance.
```{r}
#Problem 2
Auto <- read.csv(file = 'Auto.csv', header = TRUE)

#The horsepower variable was read in as a character instead of numeric for some reason so I converted it back to a numeric 
Auto$horsepower<-as.numeric(as.character(Auto$horsepower))

#problem 2a code
sapply(Auto, class)

##Problem 2b - Range
# x indicates which auto data is numeric and we estimate the rest of the problems based only on the numeric variables
x <- sapply(Auto, is.numeric)
x
sapply(Auto[, x], range, na.rm = TRUE)

##Problem 2c - means and stdv of quantitative predictors 
sapply(Auto[,x], mean, na.rm = TRUE)
sapply(Auto[,x], sd, na.rm = TRUE)

##Problem 2d
Auto1 = Auto[-c(10:85),]
sapply(Auto1[, x], range, na.rm = TRUE)
sapply(Auto1[,x], mean, na.rm = TRUE)
sapply(Auto1[,x], sd, na.rm = TRUE)
 
##Problem 2e and 2f
plot(Auto)
ggplot(data = Auto,aes( x = weight,  y= cylinders)) + geom_point()
ggplot(data = Auto,aes( x = displacement ,  y= cylinders)) + geom_point()
ggplot(data = Auto,aes( x = weight ,  y= mpg)) + geom_point()
ggplot(data = Auto,aes( x = year ,  y= mpg)) + geom_point()
ggplot(data = Auto,aes( x = origin ,  y= mpg)) + geom_point()
ggplot(data = Auto,aes( x = acceleration ,  y= mpg)) + geom_point()
```
## Problem 2
a) Name is qualitative. The rest of the variables are quantitative

e)From the overall plots and the mini plots we can say that weight and cylinder are well correlated. This is also true for cylinders and displacement. There is a negative correlation between weight and mpg.
f) Based on some of the mpg plots above, year, orgin and acceleration might be useful in predicting mpg. 
```{r}
##Problem 3
#Problem 3a
library(MASS)
? Boston
nrow(Boston)
ncol(Boston)

#Problem 3b
pairs(Boston)
#Problem 3c
plot(data = Boston, crim~age)
plot(data = Boston, crim~tax)

#Problem 3d
par(mfrow = c(2,2))
hist(Boston$crim, breaks = "FD")
hist(Boston$crim, breaks = 40)
hist(Boston$tax, breaks = "FD")
hist(Boston$ptratio, breaks = "FD")

#Problem 3e
summary(Boston$chas == 1)

#Problem 3f
median(Boston$ptratio)

#Problem 3g
which.min(Boston$medv)

#Problem 3h
summary(Boston$rm > 7)
summary(Boston$rm > 8)

```
## Problem 3
a)  The Boston data frame contains 506 rows and 14 columns.
b) chas is a binary variable. rad and tax contains lots of constant variables.
c) Based on the pairwise plot and the plot above and the individual plots of crim and two other variables above, we can see that there exists a relationship between crim  and some of the other predictors like age , tax and so on.
d) The Boston$crim historgram has a long tale. There are outliers for the Boston$tax histogram and the  Boston$ptratio plot has no real outliers.
e)    Mode   FALSE    TRUE 
logical     471      35 
f) Median = 19.05
g) suburb #399
h) for 7 rooms per dwelling we have
   Mode   FALSE    TRUE 
logical     442      64 
 for 8 rooms per dwelling we have
   Mode   FALSE    TRUE 
logical     493      13 
 




