---
title: "R Notebook"
output:   
  html_document:
    df_print: paged
---

ISyE 603
Course Project
Group 1


1. Load Libraries and Set Directory for files with stock price data.

```{r setup, include=TRUE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library('readxl')
library(forecast)
library(tensorflow)
library(keras)
library(dplyr)
library(ISLR)
library("caret")
library("pROC")
library(matrixStats)
library(fastDummies)
library(tidyverse)
library(xts)
library(tibbletime)
library(tseries)

setwd("C:/Users/mnaje/OneDrive/Documents/ISyE 603/Course Project")


```


2. Load data for each of the three stocks. 

```{r}

data_Apple = read.csv('AAPL.csv', header = TRUE)
data_Netflix = read.csv('NETFLIX.csv', header = TRUE)
data_Tesla = read.csv('TSLA.csv', header = TRUE)

```


3. Convert the data for each of the three stocks into a weekly time series. Use data starting with the first trading session of 2014.

Reference: https://community.rstudio.com/t/problem-with-converting-daily-stock-series-into-monthly-ones/3923
```{r}

#Apple
dates1 = as.Date(data_Apple$Date)
data_Apple$Date = dates1

data_Apple_Filter = filter(data_Apple, Date >= as.Date("2014-01-02"), Date <= as.Date("2021-12-31"))

data_Apple_Weekly = data_Apple_Filter %>%
  as_tbl_time(Date) %>%
  mutate(Date = collapse_index(Date,"weekly")) %>%
  group_by(Date) %>%
  summarize(
    Open = first(Open),
    High = max(High),
    Low = min(Low),
    Close = last(Close),
    Adj.Close = last(Adj.Close),
    Volume = sum(Volume)
  )

apple_ts = ts(data_Apple_Weekly$Close, start = c(2014,1), end = c(2021, 52), freq=52)


#Netflix
dates2 = as.Date(data_Netflix$Date, format = "%m/%d/%Y")
data_Netflix$Date = dates2

data_Netflix_Filter = filter(data_Netflix, Date >= as.Date("2014-01-02"), Date <= as.Date("2021-12-31"))

data_Netflix_Weekly = data_Netflix_Filter %>%
  as_tbl_time(Date) %>%
  mutate(Date = collapse_index(Date,"weekly")) %>%
  group_by(Date) %>%
  summarize(
    High = max(High),
    Low = min(Low),
    Open = first(Open),
    Close = last(Close),
    Volume = sum(Volume),
    Adj.Close = last(Adj.Close)
  )

netflix_ts = ts(data_Netflix_Weekly$Close, start = c(2014,1), end = c(2021, 52), freq=52)


#Tesla
dates3 = as.Date(data_Tesla$Date)
data_Tesla$Date = dates3

data_Tesla_Filter = filter(data_Tesla, Date >= as.Date("2014-01-02"), Date <= as.Date("2021-12-31"))

data_Tesla_Weekly = data_Tesla_Filter %>%
  as_tbl_time(Date) %>%
  mutate(Date = collapse_index(Date,"weekly")) %>%
  group_by(Date) %>%
  summarize(
    Open = first(Open),
    High = max(High),
    Low = min(Low),
    Close = last(Close),
    Adj.Close = last(Adj.Close),
    Volume = sum(Volume)
  )

tesla_ts = ts(data_Tesla_Weekly$Close, start = c(2014,1), end = c(2021, 52), freq=52)

```


4. Plot the time series decomposition for each of the three stocks, and remove any trends apparent from the time series decomposition using various types of differencing. Plot the autocorrelation and partial autocorrelation for the final differenced series. 

```{r}

#Apple
apple_decomp = decompose(apple_ts)
plot(apple_decomp)

```


```{r}

#Apple

plot(apple_ts,type='l',xlab='Time',ylab='Weekly Closing Price
Price for Apple Stock (US$)',main='Weekly Closing Stock Price of Apple')

apple_data_d1=diff(data_Apple_Weekly$Close)

apple_data_d2 = diff(apple_data_d1)

apple_data_log=diff(log(data_Apple_Weekly$Close))

apple_data_sd=diff(sqrt(data_Apple_Weekly$Close))

#par(mfrow=c(4,1))
plot(apple_data_d1,type='l',xlab='Time',ylab='Difference', main='First Degree Differencing on Raw Data')
plot(apple_data_d2,type='l',xlab='Time',ylab='Difference', main='Second Degree Differencing on Raw Data')
plot(apple_data_log,type='l',xlab='Time',ylab='Difference',main='First 
Degree Differencing on Logged Data')
plot(apple_data_sd,type='l',xlab='Time',ylab='Difference',main='First 
Degree Differencing on Square-root Data')

#par(mfrow=c(2,1))
acf(apple_data_d2,main='Autocorrelation Function of the First Degree Differencing on the Log of the Data')
pacf(apple_data_d2,main='Partial Autocorrelation Function of the First Degree Differencing on the Log of the Data')

#Augmented Dickey-Fuller Test on the Original Data
adf.test(apple_ts)

#Augmented Dickey-Fuller Test on the First Degree Differencing on the Log of the Data
adf.test(apple_data_d2)


```


```{r}

#Netflix
netflix_decomp = decompose(netflix_ts)
plot(netflix_decomp)

```


```{r}

#Nexflix

plot(netflix_ts,type='l',xlab='Time',ylab='Weekly Closing Price
Price for Netflix Stock (US$)',main='Weekly Closing Stock Price of Netflix')

netflix_data_d1=diff(data_Netflix_Weekly$Close)

netflix_data_d2=diff(netflix_data_d1)

netflix_data_log=diff(log(data_Netflix_Weekly$Close))

netflix_data_sd=diff(sqrt(data_Netflix_Weekly$Close))

#par(mfrow=c(4,1))
plot(netflix_data_d1,type='l',xlab='Time',ylab='Difference', main='First Degree Differencing on Raw Data')
plot(netflix_data_d2,type='l',xlab='Time',ylab='Difference', main='Second Degree Differencing on Raw Data')
plot(netflix_data_log,type='l',xlab='Time',ylab='Difference',main='First 
Degree Differencing on Logged Data')
plot(netflix_data_sd,type='l',xlab='Time',ylab='Difference',main='First 
Degree Differencing on Square-root Data')

#par(mfrow=c(2,1))
acf(netflix_data_log,main='Autocorrelation Function of the First Differences')
pacf(netflix_data_log,main='Partial Autocorrelation Function of the First Differences')

#Augmented Dickey-Fuller Test on the Original Data
adf.test(netflix_ts)

#Augmented Dickey-Fuller Test on the First Degree Differencing on the Log of the Data
adf.test(netflix_data_log)

```


```{r}
#Tesla

tesla_decomp = decompose(tesla_ts)
plot(tesla_decomp)

```


```{r}

#Tesla

plot(tesla_ts,type='l',xlab='Time',ylab='Weekly Closing Price
Price for Tesla Stock (US$)',main='Weekly Closing Stock Price of Tesla')

tesla_data_d1=diff(data_Tesla_Weekly$Close)

tesla_data_d2=diff(tesla_data_d1)

tesla_data_log=diff(log(data_Tesla_Weekly$Close))

tesla_data_sd=diff(sqrt(data_Tesla_Weekly$Close))

#par(mfrow=c(4,1))
plot(tesla_data_d1,type='l',xlab='Time',ylab='Difference', main='First Degree Differencing on Raw Data')
plot(tesla_data_d2,type='l',xlab='Time',ylab='Difference', main='Second Degree Differencing on Raw Data')
plot(tesla_data_log,type='l',xlab='Time',ylab='Difference',main='First 
Degree Differencing on Logged Data')
plot(tesla_data_sd,type='l',xlab='Time',ylab='Difference',main='First 
Degree Differencing on Square-root Data')

#par(mfrow=c(2,1))
acf(tesla_data_log,main='Autocorrelation Function of the First Differences')
pacf(tesla_data_log,main='Partial Autocorrelation Function of the First Differences')

#Augmented Dickey-Fuller Test on the Original Data
adf.test(tesla_ts)

#Augmented Dickey-Fuller Test on the First Degree Differencing on the Log of the Data
adf.test(tesla_data_log)

```


5. Select the best model for the data using the last year (2021) as the validation set and the years 2014 through 2020 as the training set.  

```{r}
#Apple
valid_len = 52
train_len = length(apple_ts) - valid_len

apple_train = window(apple_ts, start = c(2014,1), end = c(2014,train_len))
apple_valid = window(apple_ts, start = c(2014, train_len + 1), end = c(2014, train_len + valid_len))


# TSLM
lm_apple_train = tslm(apple_train ~ trend + season)
lm_apple_train
par(mfrow = c(1,2))

acf(lm_apple_train$residuals, main = "tslm acf for Apple")
pacf(lm_apple_train$residuals, main = "tslm pcf for Apple")

# ARIMA
aa_apple_train = auto.arima(apple_train)
aa_apple_train

par(mfrow = c(1,2))
acf(aa_apple_train$residuals, main = "Arima acf for Apple")
pacf(aa_apple_train$residuals, main = "Arima pcf for Apple")

## Neural Net
## seasonal lag as 26
nn_apple_train = nnetar(apple_train)
nn_apple_train


par(mfrow = c(1,2))
acf(nn_apple_train$residuals, main = "NN acf for Apple", na.action = na.pass)
pacf(nn_apple_train$residuals, main = "NN pcf for Apple",  na.action = na.pass)

# Predictions for validation period
tslm_apple_pred <- forecast(lm_apple_train,h=valid_len)
arima_apple_pred <- forecast(aa_apple_train,h=valid_len)
nn_apple_pred <- forecast(nn_apple_train,h=valid_len)

# Predictions for validation period plus one year
tslm_apple_pred_1 <- forecast(lm_apple_train,h=(valid_len+52))
arima_apple_pred_1 <- forecast(aa_apple_train,h=(valid_len+52))
nn_apple_pred_1 <- forecast(nn_apple_train,h=(valid_len+52))


par(mfrow = c(1,1))
plot(apple_train, col="black", ylab = "Stock Price (US$)")
lines(tslm_apple_pred$fitted, col="blue")
lines(arima_apple_pred$fitted, col="green")
lines(nn_apple_pred$fitted, col="red")

legend("topleft",c("Actual","TSLM", "ARIMA", "NN"),col=c("black","blue", "green", "Red"), lty = 1)

```


```{r}

x = 1/12
y = 2022 - x

plot(apple_valid, xlab = "2021", ylab = "Stock Price (US$)", col="black", main = "Forecasted Apple Stock Price for Validation Period", lwd = 2, xaxt="n")
axis(1, at=seq(2021, y, x), labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
lines(arima_apple_pred$mean, col="green")

```


```{r}

tslm_apple_pred_1
arima_apple_pred_1
nn_apple_pred_1

tslm_apple_pred_1 %>%
  autoplot()

arima_apple_pred_1 %>%
  autoplot()

nn_apple_pred_1 %>%
  autoplot()

```


```{r}
# Training Accuracy
accuracy(tslm_apple_pred$fitted,apple_train)
accuracy(arima_apple_pred$fitted,apple_train)
accuracy(nn_apple_pred$fitted,apple_train)

```


```{r}
# Pred Accuracy
accuracy(tslm_apple_pred$mean,apple_valid)
accuracy(arima_apple_pred$mean,apple_valid)
accuracy(nn_apple_pred$mean,apple_valid)
```


```{r}
# Residuals
plot(tslm_apple_pred$residuals, col="black")
lines(arima_apple_pred$residuals, col="green")
lines(nn_apple_pred$residuals, col="red")

legend("top",c("TSLM", "ARIMA", "NN"),col=c("black", "green", "Red"), lty = 1)

```


```{r}
#Multi-Step Re-estimation Using ARIMA model (use all of the data preceding each observation, both training and test data, to compute the forecasted values)

h = 1 #steps ahead
n = valid_len - h + 1
apple_fit = auto.arima(apple_train)
apple_fc = ts(numeric(n), start=2021+(h-1)/52, frequency = 52)  
for (i in 1:n)
{
  x = window(apple_ts, end = 2020.99 + (i-1)/52)
  apple_refit = Arima(x, model = apple_fit)
  apple_fc[i] = forecast(apple_refit, h = h)$mean[h]
}


accuracy(apple_fc,apple_valid)

```


```{r}
x = 1/12
y = 2022 - x

plot(apple_valid, xlab = "2021", ylab = "Stock Price (US$)", col="black", main = "Forecasted Apple Stock Price for Validation Period Using 1-Step Re-Estimation", lwd = 2, xaxt="n")
axis(1, at=seq(2021, y, x), labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
lines(apple_fc, col="green")


```
Reference: https://robjhyndman.com/hyndsight/rolling-forecasts/


```{r}

#Netflix
valid_len = 52
train_len = length(netflix_ts) - valid_len

netflix_train = window(netflix_ts, start = c(2014,1), end = c(2014,train_len))
netflix_valid = window(netflix_ts, start = c(2014, train_len + 1), end = c(2014, train_len + valid_len))


# TSLM
lm_netflix_train = tslm(netflix_train ~ trend + season)
lm_netflix_train
par(mfrow = c(1,2))

acf(lm_netflix_train$residuals, main = "tslm acf for Netflix")
pacf(lm_netflix_train$residuals, main = "tslm pcf for Netflix")

# ARIMA
aa_netflix_train = auto.arima(netflix_train)
aa_netflix_train

par(mfrow = c(1,2))
acf(aa_netflix_train$residuals, main = "Arima acf for Netflix")
pacf(aa_netflix_train$residuals, main = "Arima pcf for Netflix")

## Neural Net
## seasonal lag as 26
nn_netflix_train = nnetar(netflix_train)
nn_netflix_train


par(mfrow = c(1,2))
acf(nn_netflix_train$residuals, main = "NN acf for Netflix", na.action = na.pass)
pacf(nn_netflix_train$residuals, main = "NN pcf for Netflix",  na.action = na.pass)

# Predictions for validation period
tslm_netflix_pred <- forecast(lm_netflix_train,h=valid_len)
arima_netflix_pred <- forecast(aa_netflix_train,h=valid_len)
nn_netflix_pred <- forecast(nn_netflix_train,h=valid_len)

# Predictions for validation period plus one year
tslm_netflix_pred_1 <- forecast(lm_netflix_train,h=(valid_len+52))
arima_netflix_pred_1 <- forecast(aa_netflix_train,h=(valid_len+52))
nn_netflix_pred_1 <- forecast(nn_netflix_train,h=(valid_len+52))


par(mfrow = c(1,1))
plot(netflix_train, col="black", ylab = "Stock Price (US$)")
lines(tslm_netflix_pred$fitted, col="blue")
lines(arima_netflix_pred$fitted, col="green")
lines(nn_netflix_pred$fitted, col="red")

legend("topleft",c("Actual","TSLM", "ARIMA", "NN"),col=c("black","blue", "green", "Red"), lty = 1)

```


```{r}

x = 1/12
y = 2022 - x

plot(netflix_valid, xlab = "2021", ylab = "Stock Price (US$)", col="black", main = "Forecasted Netflix Stock Price for Validation Period", lwd = 2, xaxt="n")
axis(1, at=seq(2021, y, x), labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
lines(arima_netflix_pred$mean, col="green")

```


```{r}

tslm_netflix_pred_1
arima_netflix_pred_1
nn_netflix_pred_1

tslm_netflix_pred_1 %>%
  autoplot()

arima_netflix_pred_1 %>%
  autoplot()

nn_netflix_pred_1 %>%
  autoplot()

```


```{r}
# Training Accuracy
accuracy(tslm_netflix_pred$fitted,netflix_train)
accuracy(arima_netflix_pred$fitted,netflix_train)
accuracy(nn_netflix_pred$fitted,netflix_train)

```


```{r}
# Pred Accuracy
accuracy(tslm_netflix_pred$mean,netflix_valid)
accuracy(arima_netflix_pred$mean,netflix_valid)
accuracy(nn_netflix_pred$mean,netflix_valid)
```


```{r}
# Residuals
plot(tslm_netflix_pred$residuals, col="black")
lines(arima_netflix_pred$residuals, col="green")
lines(nn_netflix_pred$residuals, col="red")

legend("top",c("TSLM", "ARIMA", "NN"),col=c("black", "green", "Red"), lty = 1)

```


```{r}
#Multi-Step Re-estimation Using ARIMA model (use all of the data preceding each observation, both training and test data, to compute the forecasted values)

h = 1 #steps ahead
n = valid_len - h + 1
netflix_fit = auto.arima(netflix_train)
netflix_fc = ts(numeric(n), start=2021+(h-1)/52, frequency = 52)  
for (i in 1:n)
{
  x = window(netflix_ts, end = 2020.99 + (i-1)/52)
  netflix_refit = Arima(x, model = netflix_fit)
  netflix_fc[i] = forecast(netflix_refit, h = h)$mean[h]
}


accuracy(netflix_fc,netflix_valid)

```


```{r}
x = 1/12
y = 2022 - x

plot(netflix_valid, xlab = "2021", ylab = "Stock Price (US$)", col="black", main = "Forecasted Netflix Stock Price for Validation Period Using 1-Step Re-Estimation", lwd = 2, xaxt="n")
axis(1, at=seq(2021, y, x), labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
lines(netflix_fc, col="green")


```
Reference: https://robjhyndman.com/hyndsight/rolling-forecasts/


```{r}

#Tesla
valid_len = 52
train_len = length(tesla_ts) - valid_len

tesla_train = window(tesla_ts, start = c(2014,1), end = c(2014,train_len))
tesla_valid = window(tesla_ts, start = c(2014, train_len + 1), end = c(2014, train_len + valid_len))


# TSLM
lm_tesla_train = tslm(tesla_train ~ trend + season)
lm_tesla_train
par(mfrow = c(1,2))

acf(lm_tesla_train$residuals, main = "tslm acf for Tesla")
pacf(lm_tesla_train$residuals, main = "tslm pcf for Tesla")

# ARIMA
aa_tesla_train = auto.arima(tesla_train)
aa_tesla_train

par(mfrow = c(1,2))
acf(aa_tesla_train$residuals, main = "Arima acf for Tesla")
pacf(aa_tesla_train$residuals, main = "Arima pcf for Tesla")

## Neural Net
## seasonal lag as 26
nn_tesla_train = nnetar(tesla_train)
nn_tesla_train


par(mfrow = c(1,2))
acf(nn_tesla_train$residuals, main = "NN acf for Tesla", na.action = na.pass)
pacf(nn_tesla_train$residuals, main = "NN pcf for Tesla",  na.action = na.pass)

# Predictions for validation period
tslm_tesla_pred <- forecast(lm_tesla_train,h=valid_len)
arima_tesla_pred <- forecast(aa_tesla_train,h=valid_len)
nn_tesla_pred <- forecast(nn_tesla_train,h=valid_len)

# Predictions for validation period plus one year
tslm_tesla_pred_1 <- forecast(lm_tesla_train,h=(valid_len+52))
arima_tesla_pred_1 <- forecast(aa_tesla_train,h=(valid_len+52))
nn_tesla_pred_1 <- forecast(nn_tesla_train,h=(valid_len+52))


par(mfrow = c(1,1))
plot(tesla_train, col="black", ylab = "Stock Price (US$)")
lines(tslm_tesla_pred$fitted, col="blue")
lines(arima_tesla_pred$fitted, col="green")
lines(nn_tesla_pred$fitted, col="red")

legend("topleft",c("Actual","TSLM", "ARIMA", "NN"),col=c("black","blue", "green", "Red"), lty = 1)

```


```{r}
x = 1/12
y = 2022 - x

plot(tesla_valid, xlab = "2021", ylab = "Stock Price (US$)", col="black", main = "Forecasted Tesla Stock Price for Validation Period", lwd = 2, xaxt="n")
axis(1, at=seq(2021, y, x), labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
lines(arima_tesla_pred$mean, col="green")
```


```{r}

tslm_tesla_pred_1
arima_tesla_pred_1
nn_tesla_pred_1

tslm_tesla_pred_1 %>%
  autoplot()

arima_tesla_pred_1 %>%
  autoplot()

nn_tesla_pred_1 %>%
  autoplot()

```


```{r}
# Training Accuracy
accuracy(tslm_tesla_pred$fitted,tesla_train)
accuracy(arima_tesla_pred$fitted,tesla_train)
accuracy(nn_tesla_pred$fitted,tesla_train)

```


```{r}
# Pred Accuracy
accuracy(tslm_tesla_pred$mean,tesla_valid)
accuracy(arima_tesla_pred$mean,tesla_valid)
accuracy(nn_tesla_pred$mean,tesla_valid)
```


```{r}
# Residuals
plot(tslm_tesla_pred$residuals, col="black")
lines(arima_tesla_pred$residuals, col="green")
lines(nn_tesla_pred$residuals, col="red")

legend("top",c("TSLM", "ARIMA", "NN"),col=c("black", "green", "Red"), lty = 1)

```


```{r}
#Multi-Step Re-estimation Using ARIMA model (use all of the data preceding each observation, both training and test data, to compute the forecasted values)

h = 1 #steps ahead
n = valid_len - h + 1
tesla_fit = auto.arima(tesla_train)
tesla_fc = ts(numeric(n), start=2021+(h-1)/52, frequency = 52)  
for (i in 1:n)
{
  x = window(tesla_ts, end = 2020.99 + (i-1)/52)
  tesla_refit = Arima(x, model = tesla_fit)
  tesla_fc[i] = forecast(tesla_refit, h = h)$mean[h]
}


accuracy(tesla_fc,tesla_valid)

```


```{r}
x = 1/12
y = 2022 - x

plot(tesla_valid, xlab = "2021", ylab = "Stock Price (US$)", col="black", main = "Forecasted Tesla Stock Price for Validation Period Using 1-Step Re-Estimation", lwd = 2, xaxt="n")
axis(1, at=seq(2021, y, x), labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
lines(tesla_fc, col="green")


```
Reference: https://robjhyndman.com/hyndsight/rolling-forecasts/




