---
Name: "Zanub Hassan"
title: "IsyE 603 HW2"
output: html_notebook
---


```{r}
#Problem 1 - Estimating sample mean and sample standard dev
#Creating the data
v <- c(96, 102, 104, 108, 126, 128, 150, 156)
v
#Sample mean
mean(v)
#Sample stdev
sd(v)
```
```{r}
#problem 2 - Plot histogram from csv data
df<- read.csv(file = 'h2p3.csv', header = FALSE)
hist(df$V1, breaks = 10)
```
##Problem 2 written answer
The histogram has a bell shape which resembles a normal distribution.
```{r}
#Problem 3- probability
integrand1<-function(x){4*(x-11.75)}
#Note we also integrate from -inf to 12 but that equals 0 beacuse f(x) = 0 for anything that is not in the given boundary for f(x)
#firstfunc<-integrate(integrand1,lower = -Inf, upper = 12) = 0
func<-integrate(integrand1,lower = 11.75, upper = 12)
func
```
## Problem 4
probability that variable is larger than 4.5 or smaller than -2.5
mean = 1
standard dev = 1

z =(x - mean)/stdev
#using the formula above

P(x > 4.5) = (4.5-1)/1 = 3.5
So we look for the value of 3.0 under 5 in the z norm table
= 0.99886
The z score table normally returns for x less than a value but this time around the function is asking for x > 4.5 therefore to get the  probability we subtract the value we got from the table from unity
= 1-0.99886 = 0.00114

 P(X < -2.5) = (-2.5 -1)/1 = -3.5
 so we look for the value of -3.0 under 5 in the z norm table
 = 0.00114
  Therefore the probability that variable is larger than 4.5 or less than -2.5 is 0.00114.
  
```{r}
#Problem 5a - autocorrelation for lag 1, 2, 3
library(tseries)
x <- c(482, 490, 491, 495, 499, 499, 503, 507, 509, 510)
acf(x,lag.max = 3, pl = FALSE)
#Problem 5b
acf(x,lag.max = 3)
```
## Problem  5b - written answer
From the plot above, since the autocorrelation spike at 1 rises above the dash line, we can say it is significant. However, the autocorrelation is not significant for 2 and 3 since the spikes are not above the dashed lines.

## Problem 6 - fpp3 book chapter 2 problem 10
Matching the time series plot to the autocorrelation plot
(1,B),  (3,D), (2,A), (4,C)

```{r}
#Problem 7a - computing a time series plot
library(tidyverse)
library(dplyr)
df1<- read.csv(file = 'SouvenirSales.csv', header = TRUE)
df1_ts=ts(df1$Sales,start=c(1995,1), frequency=12)
plot(df1_ts, xlab ="Time", ylab = "Sales", main ="Time series plot of Sales vs Time")

#Problem 7b -change  xy into a log scale
plot.ts(df1_ts,xlab ="Time", ylab = "Sales", log = "xy", main = "logarithmic Time series Plot ")


```
## Problem 7b and C - written
7b) Plot b is more linear (logarithmic plot)

7c)There is an upward trend in the data based on both plots
  
```{r}
#Problem 8
library(fpp3)
library(ggplot2)
#Box cox transformation for Tobacco in aus_production data of fpp3
#Step 1 is to find lambda
lambda <- aus_production %>%
  features(Tobacco, features = guerrero) %>%
pull(lambda_guerrero)
lambda

aus_production %>%
  autoplot(box_cox(Tobacco, lambda)) 
```
```{r}
#Problem 9 - Conducting STL decomposition for Canadian Gas data in fpp3
library(fpp3)
canadian_gas %>%
  model(
    STL(Volume ~ trend(window = 7) +    #Volume parameter comes from canadian_gas
          season(window = "periodic"),
        robust = TRUE)) %>%
  components() %>%
  autoplot()

```

 

